A Few Easy Pieces: \\ How to enhance your relationship with your future self.

### 2011-03-16 workflow.Rnw
Took into account Kieran's comments

The first author convenes a three way video conference with the other
collaborators while commuting home in his flying car. The group must
go back to the analyses. Which ones? The ones from 2011? Or 2012? Or
2018? Where are the files? The next day, one member of the group finds
some of the files. Now the checks and re-analyses should be
easy. Right? The student-now-faculty member is the same person and
should remember the reason for those bits of code (or at least should
                                                   remember which series of mouse clicks were used to produce the numbers
                                                   for that crucial table --- and those mouse clicks should not be time
                                                   consuming to redo exactly as they were done in 2011, or was it
                                                   2015?). And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL
stores files and the way that your machine reads and writes them is
the same, right --- since Windows and Mac OS X have always existed and
will always continue to exist more or less as they currently exist? And
the group knows exactly which bit of code produced which table and
which figure, right? And they wrote their code following Nagler's
Maxims \citet{nagler1995coding} and being following King's Replication
Standard \citet{king1995replication} right?

@@

All tables and graphs should be generated from code, not clicking
(let alone copying and pasting). If I wanted to re-create the figure
you created but with red lines, and only for people in the South, I
should be able to do so with just a few edits to the code.

@@

The Sweave system is even nicer because you can have a LaTeX file with
R code right inside of it! I showed Sweave with plotting in
Figure~\ref{fig:giniprot}, those who view the code for this essay will
see how the Table~\ref{tab:protest} was also generated directly from a
regression object.

@@

Right now it is hard to beat Sweave as a system for enhancing the
collaboration of research groups doing quantitative work in political
science. Something will make our lives easier than Sweave (or
                                                           odfWeave) in the future. Then we'll change.

@@

\section{Enabling revision to prevent clobbering.}

Group work requires version control.\footnote{See X for more
  discussion of what version control is.} Many people are familiar
with the ``Track Changes'' feature in modern WYSIWYG word processors
or the fact that Dropbox allows one to recover previous versions of
files. These are both kinds of version control. When collaborating
with yourself or others, it is useful to see what has changed, to feel
free to experiment and then to dump the experiment in favor of
previous work, to have multiple ``releases'' of the same document (one
to MPSA, one to APSR, one to your parents) without requiring that
spawn many copies of the same document and risk confusion and
clobbering. Clobbering is what happens when your future self or your
other collaborators saves an old version of a file over a new version
--- erasing good work by accident.

Of course if you rely on Dropbox or ``track changes'' for version
control, you must communicate with other folks in your group before
you edit existing files. Only one of you can edit and save a given
file at a time. This prevents your work (or your colleagues work) from
getting lost when you both try to save the same file on top of each
other. If you find that you are needing to work on the same files at
the same time, then you should work on establishing your own shared
version control system. Free options include launchpad, github,
sourceforge for open source projects (i.e. papers you are writing
which you are happy to share with others). Each of those services
include paid versions too. One may also use Dropbox as a kind of
server for version control: checking out files from the Dropbox
directory into a local working directory. (Notice that this is
different from directly working on files within your Dropbox managed
directories.)

@@

Fancy version control systems are not required, however. I suspect
that Google Docs allows a kind of version tracking and collaboration
as well.  An excellent, simple, and robust version control system is
to merely rename your files with the date and time of saving them:
thedoc.tex becomes thedoc25-12-2011-23:50.tex.  Be sure to include
year in the file names --- remember, the life of an idea is measured
in years. Also, if you use this method, spend a little extra time to
ensure that you do not clobber files when you make typos in the
file-names.

@@

 Imagine that you reported  bootstrap confidence intervals in your famous article. Now the statisticians at the UN are worrying about your bootstrap confidence interval. You'd like to evaluate your bootstrap procedure. Although nice code exists for bootstrapping linear models, no nice code exists to bootstrap the bootstrap. Of course, the code required is not complex, but since you are writing your own code you worry about getting it right. Now, 9 years after the idea, you've had lots of time to appreciate problems arising from bugs and errors in data analysis code.

@@

         Of course, this idea, like most others is not new. When large groups of programmers write code for multi-million dollar programs the question about avoiding error looms large. The idea of \href{http://en.wikipedia.org/wiki/Test-driven_development}{test driven development} and the idea that one ought to create tests of \href{http://en.wikipedia.org/wiki/Unit_testing}{small parts of one's code} arose to address such concerns. For the social scientist collaborating with her future self and/or a small group of collaborators.

>>>>

The first author convenes a three way video conference with the other
collaborators while commuting home in his flying car. The group must
go back to the analyses. Which ones? The ones from 2011? Or 2012? Or
2018? Where are the files? The next day, one member of the group finds
some of the files. Now the checks and re-analyses should be
easy. Right? The student, now professor, is the same person and
should remember the reason for those bits of code (or at least should
                                                   remember which series of mouse clicks were used to produce the numbers
                                                   for that crucial table --- and those mouse clicks should not be time
                                                   consuming to redo exactly as they were done in 2011, or was it
                                                   2015?). And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL
stores files and the way that your machine reads and writes them is
the same, right --- since Windows and Mac OS X have always existed and
will always continue to exist more or less as they currently exist? And
the group knows exactly which bit of code produced which table and
which figure, right? And they wrote their code following Nagler's
Maxims \citep{nagler1995coding} and  King's Replication
Standard \citep{king1995replication} right?

@@

All tables and graphs should be generated from code, not clicking
(let alone copying and pasting). If I wanted to re-create the figure
you created but with red lines, including a new variable or specification, I
should be able to do so with just a few edits to the code.

@@

The Sweave system is even nicer because you can have a LaTeX file with
R code right inside of it! I showed Sweave with plotting in
Figure~\ref{fig:giniprot}. Those who view the code for this essay will
see how Table~\ref{tab:protest} was also generated directly from a
regression object.

@@

Right now it is hard to beat Sweave as a system for enhancing the
collaboration of research groups doing quantitative work in political
science. Something other than than Sweave will make our lives easier
in the future. Then we'll change.

@@

\section{Enabling revision to prevent clobbering and record history.}

Group work requires version control.\footnote{See X and Y in this
  issue for more discussion of what version control is.} Many people
are familiar with the ``Track Changes'' feature in modern WYSIWYG word
processors or the fact that Dropbox allows one to recover previous
versions of files. These are both kinds of version control. When
collaborating with yourself or others, it is useful to see what has
changed, to feel free to experiment and then to dump the experiment in
favor of previous work, to have multiple ``releases'' of the same
document (one to MPSA, one to APSR, one to your parents) without
requiring that spawn many copies of the same document and risk
confusion and clobbering. Clobbering is what happens when your future
self or your other collaborators saves an old version of a file over a
new version --- erasing good work by accident.

Of course if you rely on Dropbox or ``track changes'' for version
control, you must communicate with other folks in your group before
you edit existing files. Only one of you can edit and save a given
file at a time. This prevents your work (or your colleagues work) from
getting lost when you both try to save the same file on top of each
other. If you find that you need to work on the same files at
the same time, then you should work on establishing your own shared
version control system. Free options include launchpad, github,
sourceforge for open source projects (i.e. papers you are writing
which you are happy to share with others). Each of those services
include paid versions too. One may also use Dropbox as a kind of
server for version control: checking out files from the Dropbox
directory into a local working directory. (Notice that this is
different from directly working on files within your Dropbox managed
directories.)

@@

Fancy version control systems are not required, however. I suspect
that Google Docs allows a kind of version tracking and collaboration
as well.  An excellent, simple, and robust version control system is
to merely rename your files with the date and time of saving them:
thedoc.tex becomes thedoc25-12-2011-23:50.tex.  Be sure to include
year in the file names --- remember, the life of an idea is measured
in years. Also, if you use this method, spend a little extra time to
ensure that you do not clobber files when you make typos in the
file-names. And, you will find yourself spending extra time
reconciling changes made by different collaborators by hand that
modern version control systems (particularly the distributed kind
these days) take care of quickly and easily.

@@

 Imagine that you reported bootstrap confidence intervals in your
 famous article. Now the statisticians at the UN are worrying about
 your bootstrap confidence interval. You'd like to evaluate your
 bootstrap procedure. Although nice code exists for bootstrapping
 linear models, no nice code exists to bootstrap the bootstrap. Of
 course, the code required is not complex, but since you are writing
 your own code you worry about getting it right. Now, 9 years after
 the idea, you've had lots of time to appreciate problems arising from
 bugs and errors in data analysis and code.

@@

 Of course, this idea, like most others, is not new. When large groups
 of programmers write code for multi-million dollar programs the
 question about avoiding error looms large. The idea of
 \href{http://en.wikipedia.org/wiki/Test-driven_development}{test
   driven development} and the idea that one ought to create tests of
 \href{http://en.wikipedia.org/wiki/Unit_testing}{small parts of one's
   code} arose to address such concerns. For the social scientist
 collaborating with her future self and/or a small group of
 collaborators.
